---
title: "EDA"
author: "Justin Weltz"
date: "3/30/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readxl)
```

## Reading in Data

We read in the MWM data and connectome matrices:
```{r cars}
mwm<-read.csv("data/MWM.csv")
mwmlong<-read.csv("data/MWMlong.csv")
chasssy<-read_excel("data/CHASSSYMM3AtlasLegends.xlsx")
```


```{r}
filePaths <- list.files("data/Connectomes",
                        "\\.csv$", full.names = TRUE)
connectome_matlist <- lapply(filePaths,function(x) read.csv(x,header=FALSE) %>% as.matrix())

network <- connectome_matlist[[1]]
network2 <- connectome_matlist[[2]]
```

Playing with I-graph functions
# ```{r}
# require(igraph)
# require(qgraph)
# bin_network <- apply(network,1, function(x) x > 0)
# bin_network2 <- apply(network2,1, function(x) x > 0)
#
# graph_store1 <- qgraph(bin_network)
# graph_store2 <- graph_from_adjacency_matrix(bin_network, mode = "undirected")
# sgc1 <- cluster_spinglass(graph_store2)
# qgraph(bin_network, color = sgc1$membership)
# cfg1 <- cluster_fast_greedy(graph_store2)
# plot(cfg1, graph_store2)
#
# graph_store2
# 
# 
# graph2_store1 <- qgraph(bin_network2)
# graph2_store2 <- graph_from_adjacency_matrix(bin_network2, mode = "undirected")
# sgc2 <- cluster_spinglass(graph2_store2)
# sort(sgc2$membership) - sort(sgc1$membership)
# 
# 
# #cliques(graph_store2)
# #cohesive_blocks(graph_store2)
# coreness(graph_store2)
# #centralize(graph_store2)
# hist(degree(graph_store2))
# compare(sgc1, sgc2)
# count_triangles(graph_store2)
# ?count_triangles
# count_motifs(graph_store2)
# count_isomorphisms(graph_store2, graph2_store2)
# diameter(graph2_store2)
# plot(difference(graph_store2, graph2_store2))
# diversity(graph_store2)
# ego_size(graph_store2)
# plot(compose(graph_store2, graph2_store2))
# edge_density(graph_store2)
# plot(intersection(graph_store2, graph2_store2))
# ?diameter
# #graph.knn
# #crossing
# #modularity
# #random_walks
# #Generate Random Graphs
# # random dot product graphs from latent position vector
# #subgraph_isomorphic
# #topological sorting
# #convert a general graph into a forest
# ```

## Helper Functions

Binarizing Matrix Function:
```{r}
# takes matrix (matr) and binarizes each element 
# based on a threshold (strict less than)

binarize_matrix <- function(matr, threshold) {
  
  for (i in 1:length(matr)) {
    matr[[i]] <- ifelse(matr[[i]] < threshold, 0, 1)
  }
  
  return(matr)
}

```
Normalize
```{r}
normalize <- function(x){ 
  return((x-min(x))/(max(x)-min(x)))}
```

Difference in Networks (binarized matrices)
```{r}
# returns vector of number of edge deletions and additions needed for 
# binarized matr_1 to become binarized matr_2

network_diffs <- function(matr_1,matr_2,threshold) {
  
  # frequency matrix of -1,0,1 resulting from difference of bin. matrices
  diff_freq_matr <- (binarize_matrix(matr_2,threshold) - 
    binarize_matrix(matr_1,threshold)) %>% 
    table() %>%
    as.matrix()
  # Avoid NAs
  if(is.na(diff_freq_matr[3]))
    diff_freq_matr[3]=0
  
  return(diff_freq_matr[c(1,3)])
}
```

```{r}
network_diffs <- function(matr_1,matr_2,threshold) {
  
  # frequency matrix of -1,0,1 resulting from difference of bin. matrices
  diff_freq_matr <- binarize_matrix(matr_2,threshold) - 
    binarize_matrix(matr_1,threshold)
  add=sum(diff_freq_matr==1)
  del=sum(diff_freq_matr==-1)
  return(c(del,add))
}
```

Remove Subdivision 8 from the dataset
```{r}
indices=(chasssy$Subdivisions_7!="8_CSF")
indices=indices[-length(indices)]
for (i in 1:length(connectome_matlist)) {
  connectome_matlist[[i]] <- connectome_matlist[[i]][indices,indices]
}
```

Indexing Matrices by Runno
```{r}

# create list of identifiers corresponding to list of connectomes
regexp <- "N[[:digit:]]{5}"
matrix_labels <- lapply(filePaths,function(x) str_extract(x,regexp)) %>% unlist()

# takes vector of runnos (string format) and finds corresponding matrices
# returns vector of matrices if they exist, otherwise returns empty matrix

find_connects <- function(runnos) {
  
  idxs <- which(matrix_labels %in% runnos)
  
  if (length(idxs) == 0) {
    print("No matches")
    return(matrix())
  }
  
  return(connectome_matlist[idxs])
}

```


Sampler Function

Uniform Priors on the Psi parameters - Beta(1,1)

```{r}

sampler <- function(iter, pis, B0, B1, data){
#Iteration and Initialization of Value
R=dim(B0)[1]
n= length(data)
Cs=rep(0,n)
Psi=matrix(0.5,2,2)
Psi_save = matrix(0,iter,4)

for(t in 1:iter){
  # draw C
  for (i in 1:n){
     #Assign membership to the cluster around B1 or B0 based on pis
    Cs[i] = sample(x =  c(1,0), size = 1, prob= c(pis[i], 1-pis[i]))
  }
  # Update Psi
  add_B0=0
  del_B0=0
  add_B1=0
  del_B1=0
  
  #For each network, if the network has been grouped around B0, then compute the additions and subtractions needed to get from B_i to B0
  #if the network has been grouped around B1, then compute the additions and subtractions needed to get from B_i to B1
  for (i in 1:n){
    if (Cs[i]==0){
      #deletions from B_i to B0
      del_B0=del_B0+network_diffs(data[[i]],B0,0.5)[1]
      #additions form B_i to B0
      add_B0=add_B0+network_diffs(data[[i]],B0,0.5)[2]
    }
    else{
      #deletions from B_i to B1
      del_B1=del_B1+network_diffs(data[[i]],B1,0.5)[1]
      #additions from B_i to B1
      add_B1=add_B1+network_diffs(data[[i]],B1,0.5)[2]
    }
  }
  # Update and sample phi_00 variable - probability of having to delete an edge to get to B0
  Psi[1,1]=rbeta(1,del_B0+1,1 + R^2*n - del_B0)
  
  #Update and sample phi_01 variable - probability of having to add an edge to get to B0
  Psi[1,2]=rbeta(1,add_B0+1,1 + R^2*n - add_B0)
  
  #Update and sample phi_10 variable - probability of having to delete an edge to get to B1 
  Psi[2,1]=rbeta(1,del_B1+1,1 + R^2*n - del_B1)
  
    #Update and sample phi_11 variable - probability of having to add an edge to get to B1
  Psi[2,2]=rbeta(1,add_B1+1,1 + R^2*n - add_B1)
  
  #save the Psi parameters
  Psi_save[t,] = c(Psi[1,1], Psi[1,2], Psi[2,1], Psi[2,2])
}

names <- c("phi_00", "phi_01", "phi_10",  "phi_11")
colnames(Psi_save)<- names
return(data.frame(Psi_save))
}
```

Create the Data Set

```{r}
n=20
R=10
#Initialize the pis - stand in for standardized memory metric
unsort_pis=runif(n)

#sort the pis so that they correspond to the networks that vary from B0 to B1
pis <- sort(unsort_pis)
data_set <- list()

#Create networks varying from including a lot of 0's to a lot of 1's
for (i in 1:n){
  data_set[[i]]=matrix(sample(c(1,0),prob = c(i/(n+1),1-i/(n+1)),R^2,replace = TRUE), nrow = R, ncol = 10)
}

#Choose the two extreme networks
B0= data_set[[1]]
B1= data_set[[n]]

#Get rid of the extreme networks from the data set
data <- data_set[-c(1, n)] 
n= n-2
pis = pis[-c(1, n)]

samp <- sampler(1000, pis, B0, B1, data)

colMeans(samp)

#COMPARISON
#Unsorted pis - no effect - we can see that phi_00 and phi_11 go way up!

pis=sample(size = length(pis),x = pis)

samp_no_effect <- sampler(1000,pis, B0, B1, data)

colMeans(samp_no_effect)
```


Let's create a data set differently and see how the alogirthm behaves:

```{r}
n=20
R=10
#Initialize the pis - stand in for standardized memory metric
unsort_pis=runif(n)

#sort the pis so that they correspond to the networks that vary from B0 to B1
pis <- sort(unsort_pis)
data_set <- list()

#Create two extreme networks
B0 <- matrix(sample(c(1,0),R^2,replace = TRUE), nrow = R, ncol = R)

B1 <- matrix(sample(c(1,0),R^2,replace = TRUE), nrow = R, ncol = R)


#Create networks as linear combinations (sort of) of these two networks
combos <- 1:n/(n+1)

for (i in 1:n){
  prob <- B0*(1-combos[i]) + B1*(combos[i])
  new_data = matrix(NA, nrow = R, ncol = R)
  for (k in 1:R){
    for (j in 1:R){
      new_data[k,j] = sample(x = c(1,0),prob = c(prob[k,j], 1-prob[k,j]), size = 1)
    }
  }
  data_set[[i]]= new_data
}





data = data_set



samp <- sampler(1000, pis, B0, B1, data)

colMeans(samp)

#COMPARISON
#Unsorted pis shouldn't have an effect, and we see that the phis are consistently greater - the clusters are more disperse!

samp_no_effect <- sampler(1000,unsort_pis, B0, B1, data)

colMeans(samp_no_effect)


```




Random Brain Networks! No effect!
```{r}
n=20
R=10
#Initialize the pis - stand in for standardized memory metric
unsort_pis=runif(n)

#sort the pis so that they correspond to the networks that vary from B0 to B1
pis <- sort(unsort_pis)
data_set <- list()

#Create two extreme networks
B0 <- matrix(sample(c(1,0),R^2,replace = TRUE), nrow = R, ncol = R)

B1 <- matrix(sample(c(1,0),R^2,replace = TRUE), nrow = R, ncol = R)


#Create networks as linear combinations (sort of) of these two networks

combos <- 1:n/(n+1)

for (i in 1:n){
  new_data =matrix(sample(c(1,0),R^2,replace = TRUE), nrow = R, ncol = R)
  data_set[[i]]= new_data
}





data = data_set



samp <- sampler(1000, pis, B0, B1, data)

colMeans(samp)


samp_no_effect <- sampler(1000,rep(1/2,n), B0, B1, data)

colMeans(samp_no_effect)

```

Choose response variables
```{r}
choose_res=function(mwmlong,matrix_labels,string){
  #Choose the average speed of  every mouse
  #String is the desired variable to choose
  result=rep(0,length(matrix_labels))
  i=1
  for (label in matrix_labels) {
    result[i]=mwmlong %>% 
    filter(grepl(label,runno,fixed = TRUE)) %>% 
    select(string) %>% 
    unlist() %>% 
    mean(na.rm=TRUE)
    i=i+1
  }
  return(result)
}
#Choose avg.sw.speed as a response and normalize
avg.sw.speed=choose_res(mwmlong = mwmlong, matrix_labels = matrix_labels, "Average.SW.Speed") %>% 
           normalize()

    

```

Test on the real data--very slow
```{r}
B0=connectome_matlist[[which.min(avg.sw.speed)]]
B1=connectome_matlist[[which.max(avg.sw.speed)]]
pis=avg.sw.speed
sampler(iter=3,pis=avg.sw.speed,B0,B1,data=connectome_matlist)
```

Try a 30 by 30 submatrix---much faster
```{r}
for (i in 1:length(connectome_matlist)) {
  data[[i]] <- connectome_matlist[[i]][1:10,1:10]
}
B0=connectome_matlist[[which.min(avg.sw.speed)]][1:10,1:10]
B1=connectome_matlist[[which.max(avg.sw.speed)]][1:10, 1:10]
pis=avg.sw.speed
sampler(iter=3000,pis=avg.sw.speed,B0,B1,data=data)
```

