---
title: "EDA"
author: "Frances Hung"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(circlize)
```


# Summary of EDA
I took the element-wise variances of both the raw and binarized connectomes. 

As expected, the binarized and raw connectome matrices have greatest-variance connections which differ in terms of brain region and hemisphere. Variance in the raw matrices measures the variance in strength of white matter connections, while variance in binarized matrices measures the variance in existence of white matter connections. Connections with greatest variance, using the raw connectome matrices, come from fewer brain regions than the connections wrt binarized matrices. Some common (between raw and binarized) destination areas of greatest-variance connections are the parasubiculum and dorsal claustrum. 

With respect to hemisphere interaction, the raw connectome matrices' top-variance connections are both between and within the right and left hemispheres. For the binarized connectomes, the top-variance connections are exclusively from the left hemisphere to the right or within the left hemisphere.

In both cases (but especially the raw connectome case), greatest-variance connections appear to be symmetric: this is evidence that connections in the brain are typically equally present in one direction as the other.

Next steps include looking at distributions of specific high-variance connections. 

## Reading in Data

We read in the MWM data and connectome matrices:
```{r cars}
mwmlong<-read.csv("../data/MWMlong.csv")
chasssy<-read.csv("../data/CHASSSYMM3AtlasLegends.csv")
```


```{r}
filePaths <- list.files("../data/Connectomes",
                        "\\.csv$", full.names = TRUE)
connectome_matlist <- lapply(filePaths,function(x) read.csv(x,header=FALSE) %>% as.matrix())
```

## Helper Functions

### Binarizing Matrix Function

```{r}
# takes matrix (matr) and binarizes each element 
# based on a threshold (strict less than)

binarize_matrix <- function(matr, threshold) {
  
  for (i in 1:length(matr)) {
    matr[[i]] <- ifelse(matr[[i]] < threshold, 0, 1)
  }
  
  return(matr)
}

```

### Difference in Networks (binarized matrices)

```{r}
# returns vector of number of edge deletions and additions needed for 
# binarized matr_1 to become binarized matr_2

network_diffs <- function(matr_1,matr_2,threshold) {
  
  # frequency matrix of -1,0,1 resulting from difference of bin. matrices
  diff_freq_matr <- binarize_matrix(matr_2,threshold) - 
    binarize_matrix(matr_1,threshold)
  
  add=sum(diff_freq_matr==1)
  del=sum(diff_freq_matr==-1)
  
  return(c(del,add))
}
```

Remove Subdivision 8 from the dataset
```{r}
indices=(chasssy$Subdivisions_7!="8_CSF")
indices=indices[-length(indices)]
for (i in 1:length(connectome_matlist)) {
  connectome_matlist[[i]] <- connectome_matlist[[i]][indices,indices]
}
```


### Indexing Matrices by Runno
```{r}

# create list of identifiers corresponding to list of connectomes
regexp <- "N[[:digit:]]{5}"
matrix_labels <- lapply(filePaths,function(x) str_extract(x,regexp)) %>% unlist()

# takes vector of runnos (string format) and finds corresponding matrices
# returns vector of matrices if they exist, otherwise returns empty matrix

find_connects <- function(runnos) {
  
  idxs <- which(matrix_labels %in% runnos)
  
  if (length(idxs) == 0) {
    print("No matches")
    return(matrix())
  }
  
  return(connectome_matlist[idxs])
}

```

## EDA

For Raw Matrices
```{r}
mean_connectome <- Reduce("+", connectome_matlist) / length(connectome_matlist)

#elt-wise: square of obs-mean difference
sds <-lapply(connectome_matlist, function(x) x-mean_connectome) %>% 
  lapply(function(x) x^2)

#sum squares of differences and divide by number of subjects
sd_connectome <-Reduce("+",sds)/(length(sds)-1)


heatmap(mean_connectome,main="Mean Connections")
heatmap(sd_connectome,main="Connection Variance")

```

```{r}
row_label <- rep(1:nrow(sd_connectome),each=ncol(sd_connectome))
col_label <- rep(1:ncol(sd_connectome),nrow(sd_connectome))
elt_sd <- sd_connectome %>% as.vector()

#join brain regions to sorted elt-wise standard_dev dataframe
sd_sort_df <- cbind(elt_sd,row_label,col_label) %>%
  as.data.frame() %>%
  arrange(desc(elt_sd)) %>%
  inner_join(.,chasssy,by=c("row_label"="index")) %>%
  inner_join(.,chasssy,by=c("col_label"="index"))


```


### Visualization of White Matter- Origins
```{r}
sd_out_bar <- ggplot(sd_sort_df %>% slice(1:90),aes(x=Structure.x,fill=elt_sd)) + 
  geom_bar(stat="count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter From Brain Region Is in Top-90 Variable Connections")

sd_out_violin <- ggplot(sd_sort_df %>% slice(1:90),aes(x=Structure.x,y=elt_sd)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter From Brain Region Is in Top-90 Variable Connections")


ggsave("plots/sd_outgoing_raw.png",plot=gridExtra::grid.arrange(sd_out_bar, sd_out_violin, ncol=1),
       width=5,height=9)
```


### Visualization of White Matter- Destinations
```{r}
sd_in_bar <- ggplot(sd_sort_df %>% slice(1:90),aes(x=Structure.y,fill=elt_sd)) + 
  geom_bar(stat="count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter To Brain Region Is in Top-90 Variable Connections")

sd_in_violin <- ggplot(sd_sort_df %>% slice(1:90),aes(x=Structure.y,y=elt_sd)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter To Brain Region Is in Top-90 Variable Connections")


ggsave("plots/sd_incoming_raw.png",plot=gridExtra::grid.arrange(sd_in_bar, sd_in_violin, ncol=1),
       width=5,height=9)
```

### Circle Diagrams: Brain Regions
```{r}
top_sd_sort <- sd_sort_df[1:10,]

png("plots/brain_region_circle_raw.png",width=700,height=700)
chordDiagram(top_sd_sort %>% select(Structure.x,Structure.y), annotationTrack = "grid", preAllocateTracks = 1)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  circos.text(quantile(xlim,0.25), ylim[1] + 1, sector.name,niceFacing = TRUE, adj = c(0, 0.5), cex=0.9)
  circos.axis(h = "top", labels.cex = 0.5, major.tick.percentage = 0.2, sector.index = sector.name, track.index = 2)
}, bg.border = NA)
title("10 Connections w/ Most Variance (Raw): Brain Region Interaction")
dev.off()
```
### Circle Diagrams: Right and Left Hemisphere
```{r}
png("plots/hemisphere_circle_raw.png",width=700,height=700)
chordDiagram(top_sd_sort %>% select(Hemisphere.x,Hemisphere.y))
title("10 Connections w/ Most Variance (Raw): Hemisphere Interaction")
dev.off()
```


## For Binarized Matrices
```{r}
bin_connectome_matlist <- lapply(connectome_matlist, function(x) binarize_matrix(x,1))
bin_mean_connectome <- Reduce("+", bin_connectome_matlist) / length(bin_connectome_matlist)

#elt-wise: square of obs-mean difference
bin_sds <-lapply(bin_connectome_matlist, function(x) x-bin_mean_connectome) %>% 
  lapply(function(x) x^2)

#sum squares of differences and divide by number of subjects
bin_sd_connectome <-Reduce("+",bin_sds)/(length(bin_sds)-1)

mean_heatmap_bin <- (heatmap(bin_mean_connectome, main="Mean Binary Connections"))
var_heatmap_bin <- heatmap(bin_sd_connectome, main="Variance of Binary Connections")

```

```{r}
row_label <- rep(1:nrow(bin_sd_connectome),each=ncol(bin_sd_connectome))
col_label <- rep(1:ncol(bin_sd_connectome),nrow(bin_sd_connectome))
elt_sd <- bin_sd_connectome %>% as.vector()

#join brain regions to sorted elt-wise standard_dev dataframe
bin_sd_sort_df <- cbind(elt_sd,row_label,col_label) %>%
  as.data.frame() %>%
  arrange(desc(elt_sd)) %>%
  inner_join(.,chasssy,by=c("row_label"="index")) %>%
  inner_join(.,chasssy,by=c("col_label"="index"))


```


```{r}
sd_out_bar_bin <- ggplot(bin_sd_sort_df %>% slice(1:10),aes(x=Structure.x,fill=elt_sd)) + 
  geom_bar(stat="count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter From Brain Region Is in Top-10 Variable Connections")

sd_out_violin_bin <- ggplot(bin_sd_sort_df %>% slice(1:10),aes(x=Structure.x,y=elt_sd)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter From Brain Region Is in Top-10 Variable Connections")


ggsave("plots/sd_outgoing_bin.png",plot=gridExtra::grid.arrange(sd_out_bar_bin, sd_out_violin_bin, ncol=1),
       width=5,height=9)
```

```{r}
sd_in_bar_bin <- ggplot(bin_sd_sort_df %>% slice(1:10),aes(x=Structure.y,fill=elt_sd)) + 
  geom_bar(stat="count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter To Brain Region Is in Top-10 Variable Connections")

sd_in_violin_bin <- ggplot(bin_sd_sort_df %>% slice(1:10),aes(x=Structure.y,y=elt_sd)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("# Times White Matter To Brain Region Is in Top-10 Variable Connections")


ggsave("plots/sd_incoming_bin.png",plot=gridExtra::grid.arrange(sd_in_bar_bin, sd_in_violin_bin, ncol=1),
       width=5,height=10)
```

### Circle Diagram: Brain Region (Binarized)
```{r}
bin_top_sd_sort <- bin_sd_sort_df[1:10,]
png("plots/brain_region_circle_bin.png",width=700,height=700)
chordDiagram(bin_top_sd_sort %>% select(Structure.x,Structure.y), annotationTrack = "grid", preAllocateTracks = 1)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  ylim = get.cell.meta.data("ylim")
  sector.name = get.cell.meta.data("sector.index")
  circos.text(quantile(xlim,0.25), ylim[1] + 1, sector.name,niceFacing = TRUE, adj = c(0, 0.5), cex=0.9)
  circos.axis(h = "top", labels.cex = 0.5, major.tick.percentage = 0.2, sector.index = sector.name, track.index = 2)
}, bg.border = NA)

title("10 Connections w/ Most Variance (Binarized): Brain Region Interaction")
dev.off()
```
## Circle Diagram: Hemispheres (Binarized)
```{r}
png("plots/hemisphere_circle_bin.png",width=700,height=700)
chordDiagram(bin_top_sd_sort %>% select(Hemisphere.x,Hemisphere.y))
title(main="10 Connections w/ Most Variance (Binarized): Hemisphere Interaction")
dev.off()
```




<!-- Gibbs Sampler -->
<!-- ```{r} -->
<!-- n=3 -->
<!-- R=10 -->
<!-- pis=runif(n) -->
<!-- data=array(sample(c(0,1),R^2*n,replace = TRUE),dim=c(R,R,n)) -->
<!-- iter=100 -->
<!-- ##Initialization -->
<!-- Cs=rep(0,n) -->
<!-- B0=matrix(0,10,10) -->
<!-- B1=matrix(1,10,10) -->
<!-- Psi=matrix(.999,2,2) -->
<!-- Cs_save=matrix(0,n,iter) -->
<!-- for(t in 1:iter){ -->
<!--   # Update C -->
<!--   for (i in 1:n) { -->
<!--     p1=Psi[2,]^network_diffs(data[,,i],B1,0.5) -->
<!--     p1=p1[1]*p1[2]*pis[i] -->
<!--     p0=Psi[1,]^network_diffs(data[,,i],B0,0.5) -->
<!--     p0=p0[1]*p0[2]*(1-pis[i]) -->
<!--     Cs[i]=sample(c(0,1),1,prob=c(p0,p1)) -->
<!--   } -->
<!--   # Update Psi -->
<!--   add_B0=0 -->
<!--   del_B0=0 -->
<!--   add_B1=0 -->
<!--   del_B1=0 -->
<!--   for (i in 1:n) { -->
<!--     if (Cs[i]==0){ -->
<!--       del_B0=del_B0+network_diffs(data[,,i],B0,0.5)[1] -->
<!--       add_B0=add_B0+network_diffs(data[,,i],B0,0.5)[2] -->
<!--     } -->
<!--     else{ -->
<!--       del_B1=del_B1+network_diffs(data[,,i],B1,0.5)[1] -->
<!--       add_B1=add_B1+network_diffs(data[,,i],B1,0.5)[2] -->
<!--     } -->
<!--   } -->
<!--   Psi[1,1]=rbeta(1,del_B0+1,1) -->
<!--   Psi[1,2]=rbeta(1,add_B0+1,1) -->
<!--   Psi[2,1]=rbeta(1,del_B1+1,1) -->
<!--   Psi[2,2]=rbeta(1,add_B1+1,1) -->
<!--   # Update B0 -->
<!--   for (j in 1:R) { -->
<!--     for (l in 1:R) { -->
<!--       p=Psi[1,]^c(sum(data[j,l,Cs==0]==1),sum(data[j,l,Cs==0]==0)) -->
<!--       B0[j,l]=sample(c(0,1),1,prob = p) -->
<!--       p=Psi[2,]^c(sum(data[j,l,Cs==1]==1),sum(data[j,l,Cs==1]==0)) -->
<!--       B1[j,l]=sample(c(0,1),1,prob = p) -->
<!--     } -->
<!--   } -->
<!--   Cs_save[,t]=Cs -->
<!-- } -->
<!-- ``` -->

